---
title: "Homework 2"
author: "Boyu Jiang"
date: "9/11/2021"
output: pdf_document
---
# Problem 2

## Part A

| My learning objectives in this class:

* Getting familiar with R programming, typesetting, and version control; 
* Mastering various data visualization tools;
* Learning data handling skills in R.

## Part B

1. Normal Distribution
\begin{align}
f(x) = \frac{1}{\sigma \sqrt{2\pi}}\ exp \left[-\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right] \mbox{, } \sigma > 0
\end{align}

2. Exponential Distribution
\begin{align}
f(x) = \frac{1}{ \beta } exp(-\frac{x}{\beta}) \mbox{, } x \ge 0 \mbox{, } \beta > 0
\end{align}

3. Cauchy Distribution
\begin{align}
f(x) = \frac{1}{\pi \sigma}\frac{1}{1+(\frac{x-\theta}{\sigma})^2}\mbox{, } \sigma > 0
\end{align}

# Problem 3

| Steps in performing Reproducible Research:

1. Before analysis: data storage and organization.
  + Storing raw data in multiple locations using multiple media;
  + Storing final data in a portable and non-proprietary format;
  + Formatting final data appropriately for analysis. (Challenge: I found it hard to process raw data when there are various formatting problems.)

2. During analysis: best coding practices.
  + Making code clean, readable, and appropriately formatted;
  + Commenting code thoroughly;
  + Inviting at least one collaborator to review data and code;
  + Documenting all software versions and computing environments.

3. After analysis: finalizing results and sharing.
  + Giving explicit instructions on locating data, metadata, and code in the manuscript; (Challenge: although my manuscript is full of figure, table, and formula, readers or reviewers think it is indistinct.)
  + Sharing data, metadata, and code at a permanent site.

# Problem 4

```{r}
library(data.table)
covid_raw <- fread("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")
us <- covid_raw[covid_raw$countriesAndTerritories == 'United_States_of_America',]
us_filtered <- us[us$month %in% c(6:7),]
us_filtered$index <- rev(1:dim(us_filtered)[1]) 
fit<-lm(`Cumulative_number_for_14_days_of_COVID-19_cases_per_100000`~index, 
        data=us_filtered)
```


## Part A

1. 
```{r}
library(knitr)
kable(summary(us_filtered[, c('cases','deaths','popData2019',
                      'Cumulative_number_for_14_days_of_COVID-19_cases_per_100000')]),
      format = 'simple',align='l',col.names=c('cases','deaths','popData2019',
                                              'Cumul.COVID19.Cases for 14 days per 100k'))

library(plyr)
kable(count(us_filtered, vars=c('month','year','countriesAndTerritories','geoId',
                       'countryterritoryCode','continentExp')), align='l')
```

Since the time period was limited from June to July, there are 61 time points in us_filtered data. From these two tables, no missing value exists.

2. 
```{r}
library(stargazer)
stargazer(fit, title= "Fit Results", align=TRUE, type = 'text')
```

## Part B

```{r}
library(broom)
fit.diags <- broom::augment(fit)
plot(fit, 1)
plot(fit, 2)
plot(fit, 3)
plot(fit, 5)
```

## Part C

```{r}
acf(fit.diags$.resid,type = "correlation")
```

# Problem 5

```{r}
par(mfrow=c(2,2),mar = c(2,2,1.5,0.5))
plot(fit, 1)
plot(fit, 2)
plot(fit, 3)
plot(fit, 5)
```

